---
title: "R Notebook"
output: html_notebook
---
Loading libraries and reading dataset

```{r}
library(dplyr) 
library(ggplot2)
data = read.csv("compas-scores-two-years.csv")

```


Renaming Columns

```{r}
head(data) #Checking column titles and variable type
str(data)

#Renaming column relevant names 
data = data %>% rename(Sex = sex,
                Age = age,
                Race = race,
                Is_recid = is_recid,
                Recidivism_score = decile_score.1,
                Score_text = score_text,
                Violence_score = v_decile_score,
                Priors = priors_count.1)
```


Selecting relevant columns for analysis
```{r}
data = data %>% select(Sex, Age, Race, Is_recid,Recidivism_score,Violence_score, Priors, Score_text)
```


Data Visualization

- Making graph of each races prevalence in the dataset
```{r}
str(data)
unique(data$Race) #looking at different races in dataframe

data %>% group_by(Race) %>%
  count(Race)  %>%
  ggplot(aes(x = reorder(Race, -n), y = n)) + 
  geom_col() + 
  theme_minimal() + 
  labs(title = "Representation of each Race in the Dataset", 
       x = "Race", 
       y = "Number of Individuals")

```

- Making histogram of the distribution of recidivism scores

```{r}
ggplot(data, aes(x = Recidivism_score)) + 
  geom_histogram(binwidth = 1, fill = "grey", color = "black") +
  labs(title = "Distribution of Scores", x = "Score", y = "Frequency") +
  theme_minimal()

```

- Making column chart of score distribution filled in for each race

```{r}
data %>% group_by(Race) %>%
  count(Recidivism_score) %>%
  ggplot(aes(x = Recidivism_score, y = n, fill = Race)) + 
  geom_col() +
  theme_minimal() +
  labs(title = "Distribution of Scores by Race",
       x = "Recidivism Score",
       y = "Number of Individuals")
```

- Making box plot of score distribution and removing Asian and Native American because of too few data points in set

```{r}
data %>% filter(Race != 'Native American') %>%
  filter(Race != 'Asian') %>%
  ggplot(aes(x = Race, y = Recidivism_score)) +
  geom_boxplot() +
  labs(title = "Distribution of Recidivism Scores by Race",
       x = "Race",
       y = "Recidivism Score") +
  theme_minimal()

```



